---
title: "Star Wars Character Sheet"
output:
  html_document: default
  pdf_document: default
params:
  sw_char_id: '2'
---

```{r eval=FALSE, include=FALSE}
#for demo only

#demo_x <- 5


#resp_to_df(get_person_resp(demo_x))



```








```{r include=FALSE}
library(httr)
library(jsonlite)
library(magrittr)
library(tidyverse)
library(shiny)

```


```{r include=FALSE}
base_url <- "http://swapi.dev/api/"

#endpoint urls
people_url <- paste0(base_url, 'people/') 
films_url <- paste0(base_url, 'films/')
planets_url <- paste0(base_url, 'planets/')

```







```{r helper functions, include=FALSE}

#Helper Functions 
#########################

# Create url for API resource from a character id
get_person_url <- function(id){
  person_url = paste0(people_url, id, '/')
  return(person_url)
}

# Return the response from an API GET request to person endpoint 
get_person_resp <- function(id){
  resp = GET(get_person_url(id))
  warn_for_status(resp, task = 'get person')
  return(resp)
}



# Restructure the json response from an endpoint to a dataframe of key value pairs
resp_to_df <- function (r){
  df= httr::content(r) %>% 
    unlist() %>% 
    enframe()
  
  return (df)
}



# restructure the dataframe of key value pairs into a 'tidy' dataframe
respdf_to_person_df <- function(x ){
  char_df <- x %>% 
    #filter(name %in% char_col_names) %>%
    spread(name, value) #%>%
    #mutate(id = str_remove(str_remove(x$url,api_url), '/'))
  
  return(char_df)
  
}



# function takes a url and returns the {id} value from that url
##
#   df      = dataframe with url 
#   col     = column in df above with url
#   api_url = base of the url which needs to be removed
#   id_name = name of new column which will store the {id}
id_from_col <- function(df,col, api_url, id_name){
  new_col = id_name
  
  id_df =  df %>%
    mutate(id1 = str_remove(df[,col], api_url)) %>% 
    mutate(id1 = str_remove(id1, '/')) %>% 
    rename(!!new_col := id1)
  
  return(id_df)
}


# function that returns the name of a star wars planet 
#    when given the url of that planet
#
# planetUrl = url of planet whose name will be returned


get_planet_name<- function(planetUrl){
  resp = GET(planetUrl)
  warn_for_status(resp, task = 'get planet name')
  
  df= httr::content(resp) %>% 
    unlist() %>% 
    enframe() %>% 
    filter(name =='name')
  
  return(df[,2][[1]])
  
}

# Function that returns the title of a film
#  when given the url of that film
#
# filmUrl = url of film

get_firstFilm_title<- function(filmUrl){
  resp = GET(filmUrl)
  warn_for_status(resp, task = 'get film title')
  
  df= httr::content(resp) %>% 
    unlist() %>% 
    enframe() %>% 
    filter(name =='title')
  
  return(df[,2][[1]])
  
  
  
  
}


# Function that returns the director of a film
#  when given the url of that film
#
# filmUrl = url of film

get_Film_director<- function(filmUrl){
  resp = GET(filmUrl)
  warn_for_status(resp, task = 'get film title')
  
  df= httr::content(resp) %>% 
    unlist() %>% 
    enframe() %>% 
    filter(name =='director')
  
  return(df[,2][[1]])
  
  
  
  
}

```




```{r Build Character dataframe, include=FALSE}
#create character dataframe to store values for parameterized report

char_df<- resp_to_df(get_person_resp(params$sw_char_id))


# tidy up the dataframe to place each variable in it's own column
# and add columns for character id and home planet id

 char_df<- char_df %>% 
  respdf_to_person_df() %>% 
   id_from_col(col = 'url', api_url =  people_url, id_name =  'char_id') %>% 
   id_from_col( 'homeworld', planets_url, 'home_id') 
 
 # add column for name of the home planet
 char_df <- char_df %>% 
   mutate(homeworld_name = get_planet_name(char_df[,"homeworld"][[1]]))
 
 
# as part of the unlisting function for df, single value items are kept
 # while nested lists have a list value suffixed to them. 
 
 
 # this code suffixes the value of 1 to the films column to standardize
  if ('films' %in% colnames(char_df)){

   char_df <- rename(char_df, films1 = films)
 }

 # Adds columns:
 #  title_first_film     = title of first film character appear
 #  director_first_film  = name of director of first film
 #  num_vehicles_piloted = number of vehicles character pilots 
 # 
   char_df <- char_df %>%
   mutate(title_first_film = get_firstFilm_title(char_df$films1[1])) %>%
   mutate(director_first_film = get_Film_director(char_df$films1[1])) %>%
   mutate(num_vehicles_piloted = ncol(select(char_df, starts_with(c('veh','star')))))


   
# Converts the string "n/a" in gender column to 
#   "non- gendered character"
   if( char_df$gender[1] == "n/a"){
   char_df$gender[1] = "non-gendered character"
 } 
   
```





# `r char_df$name[1]`
## Basic Facts
Born: `r char_df$birth_year[1]`  
Home Planet: `r char_df$homeworld_name[1]`  
Gender: `r char_df$gender[1]`  
Hair Color: `r char_df$hair_color[1]`  
Eye Color: `r char_df$eye_color[1]`  
Height: `r char_df$height[1]` cm  
Mass: `r char_df$mass[1]` kg  

## About Character
`r char_df$name[1]`  is a fictional character in the Star Wars Universe from the homeworld of `r char_df$homeworld_name[1]` . The character made his first appearance in the Star Wars Movie, ' `r char_df$title_first_film`' which was directed by `r char_df$director_first_film[1]`. This character is a `r char_df$gender[1]` born in the year `r char_df$birth_year[1]` with a listed height of `r char_df$height[1]` cm and a mass of `r char_df$mass[1]` kg.

Though out the Star Wars films, `r char_df$name[1]` is seen piloting a total of `r char_df$num_vehicles_piloted[1]` vehicles.

